Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-03-01 paper/main.tex
first part of Lemma 3 proven

A \emph{path} from a state $p$ to a state $q$ is a finite sequence of
adjacent transitions of the form
$(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
is called a nontrivial \emph{strongly connected component (SCC)} if
for each $p,q\in S$ there is a path from $p$ to $q$. An SCC S is
\emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
the rest of this paper, SCC always refers to a maximal SCC. Given an
SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
denotes the set of all transitions between states in $S$. Further, for
each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
marked with $m$, i.e., $\delta_m=\{t\in\delta\mid m\in
\mks(t)\}$. \todo{The sets $\delta_m$ are called \emph{acceptance
    sets}. Given an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we
  denote the set $\delta_{S,m}=\delta_S\cap\delta_m$.} A set of
transitions $T\subseteq\delta$ is called a \emph{cycle} is there
exists a path from a state $p$ to the same state containing all
transitions of $T$ and no transition outside $T$. Finally, we assume
that each TELA $\mA$ contains only states $q$ that are reachable from
the initial state $q_I$ (i.e., $q=q_I$ or there is a path from $q_I$
to $q$) as states that are not reachable from $q_I$ can be eliminated
without any impact on $L(\mA)$.

@@

\subsection{Strict version $\mcycle_3(\vec{e})$}
Before we give the definition of $\mcycle_3(\vec{e})$, we prove that
cycles can be characterised in the following way.
\begin{lemma}
  A set $T\subseteq\delta$ is a cycle if for each set of
  states $S\subseteq Q$ it holds that
  \[
    \begin{split}
    \textrm{if $T$ contains a transition leading from a state in $S$ to a state outside $S$,~~~~}\\[-.5ex]
    \textrm{then it also contains a transition leading from a state outside $S$ to a state in $S$.}
    \end{split}
    \tag{$3$}\label{eq:lem3} 
  \]  
\end{lemma}
\begin{proof}
  \todo{Ted to musim dokazat.}
\end{proof}

>>>>

A \emph{path} from a state $p$ to a state $q$ is a finite sequence of
adjacent transitions of the form
$(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
is called a nontrivial \emph{strongly connected component (SCC)} if
for each $p,q\in S$ there is a path from $p$ to $q$. An SCC S is
\emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
the rest of this paper, SCC always refers to a maximal SCC. Given an
SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
denotes the set of all transitions between states in $S$. Further, for
each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
marked with $m$, i.e., $\delta_m=\{t\in\delta\mid m\in
\mks(t)\}$. \todo{The sets $\delta_m$ are called \emph{acceptance
    sets}. Given an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we
  denote the set $\delta_{S,m}=\delta_S\cap\delta_m$.} A set of
transitions $T\subseteq\delta$ is called a \emph{cycle} is there
exists a path from a state $p$ to the same state containing all
transitions of $T$ (potentially multiple times) and no transition
outside $T$. Finally, we assume that each TELA $\mA$ contains only
states $q$ that are reachable from the initial state $q_I$ (i.e.,
$q=q_I$ or there is a path from $q_I$ to $q$) as states that are not
reachable from $q_I$ can be eliminated without any impact on $L(\mA)$.

@@

\subsection{Strict version $\mcycle_3(\vec{e})$}
Before we give the definition of $\mcycle_3(\vec{e})$, we prove that
cycles can be characterised in the following way.
\begin{lemma}
  A set $T\subseteq\delta$ is a cycle if and only if for each set of
  states $S\subseteq Q$ it holds that
  \begin{equation}
    \begin{minipage}{0.85\linewidth}
      \centering
      if $T$ contains a transition leading from a state in $S$ to a state outside $S$,\\
      then it also contains a transition leading from a state outside $S$ to a state in $S$.
    \end{minipage}
    % \begin{split}
    % \textrm{if $T$ contains a transition leading from a state in $S$ to a state outside $S$,}\\[-.5ex]
    % \textrm{then it also contains a transition leading from a state outside $S$ to a state in $S$.}
    % \end{split}
    \tag{$3$}\label{eq:lem3} 
  \end{equation}
\end{lemma}
\begin{proof}
  We first prove direction ``$\Longrightarrow$''. Let $T$ be a cycle
  and $S\subseteq Q$ be a set of states such that there exists a
  transition $t\in T$ leading from a state $p\in S$ to a state
  $q\not\in S$. The definition of a cycle says that there exists a
  path $t_1t_2\ldots t_n\in T^+$ from a state $p'$ back to $p'$ using
  all transitions of $T$. Hence, $t=t_i$ for some $1\le i\le n$. Then
  $t_{i+1}\ldots t_nt_1t_2\ldots t_{i-1}\in T^+$ is a path leading
  from the state $q\not\in S$ to the state $p\in S$. This path has to
  contain a transition $t'\in T$ leading from a state outside $S$ to a
  state in $S$.

  Now we prove the opposite direction ``$\Longleftarrow$''.

\end{proof}
